---
title: "ConsumerUseData"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


Import data sheets from sheds-ht package repo
```{r}
setwd("C:/Users/ddawso01/OneDrive - Environmental Protection Agency (EPA)/Profile/Documents/Repos/SHEDSDevel/inst/extdata")
chem=read.csv("source_chem_products.csv")
vars=read.csv("source_vars_products.csv")
names(chem)=tolower(names(chem)) #make names the same case

```


Extract PUC's that correspond to byproduct product use for 1,4 Dioxane
```{r}
products=c(
  "surface cleaner",
  "antifreeze",
	"dish soap",
	 "dishwashing detergent",
	"laundry detergent",
	 "paints and floor lacquer",
	 "textile dye",
"spray polyurethane foam") 
#1. surface cleaner


```


Surface cleaner
```{r}
#Note that the risk evaluation specifies that surface cleaner includes all purpose and bathroom cleaners, so bathroom was included as a term 
scchem=(grepl("surface",chem$source.description ) & grepl("cleaner",chem$source.description )) | grepl("bathroom",chem$source.description ) 
scvars=(grepl("surface",vars$source.description ) & grepl("cleaner",vars$source.description )) | grepl("bathroom",vars$source.description ) 
chem[scchem,]
vars[scvars,]

scvars

```

Antifreeze

```{r}
afchem=grepl("antifreeze",chem$source.description ) 
afvars=grepl("antifreeze",vars$source.description ) 
chem[afchem,]
vars[afvars,]
````

Dishsoap

```{r}
dschem=grepl("dish soap",chem$source.description ) 
dsvars=grepl("dish soap",vars$source.description ) 
chem[dschem,]
vars[dsvars,]
```

Dishwashing detergent

```{r}
#note rinse aids contain surfactants; will include here with detergents but can be removed. 
dwchem=grepl("dishwashing",chem$source.description) & (grepl("detergent",chem$source.description) | grepl("additive",chem$source.description )) 
dwvars=grepl("dishwashing",vars$source.description) & (grepl("detergent",vars$source.description) | grepl("additive",vars$source.description )) 

chem[dwchem,]
vars[dwvars,]

```

Laundry detergent
```{r}
#Note, only liquid laundry detergent was used in the Risk Evaluation
ldchem=grepl("laundry detergent",chem$source.description ) 
ldvars=grepl("laundry detergent",vars$source.description ) 
chem[ldchem,]
vars[ldvars,]

```

Paints and Floor lacquer

```{r}
#Westat latex paint data was used for duration of use and mass of product in the risk evaluation. #Bedroom was specified as the room of use. Latex paint in SHeds is used in both paint-exterior and #paint-interior. THere are also craft paints, but I don't know if those woudl be the same. There are no #floor lacquers listed in sheds. However, there are wood stains-NOC, which includes interior #application. 

pfchem=(grepl("paint interior",chem$source.description) | 
  grepl("paint exterior",chem$source.description) | 
  grepl("paint NOC",chem$source.description) |
  grepl("paint spray",chem$source.description) |
  grepl("arts and crafts paint",chem$source.description) |
   grepl("stain NOC",chem$source.description )|
  grepl("stain exterior",chem$source.description)) 
  
pfvars=grepl("paint-interior",vars$source.description) | 
  grepl("paint-exterior",vars$source.description) | 
  grepl("paint-NOC",vars$source.description) |
  grepl("paint-spray",vars$source.description) |
  grepl("arts and crafts paint",vars$source.description) |
  grepl("stain-NOC",vars$source.description )|
  grepl("stain-exterior",vars$source.description )

#Note that this includes Body Paint and Auto paint which I don't know if we want to include, so I am setting up lines to take out those terms we we want
#Exclude terms
pfchemexclude=grepl("body paint NOC",chem$source.description ) | grepl("auto paint NOC",chem$source.description )  

pfvarsexclude=grepl("body paint-NOC",vars$source.description ) | grepl("auto paint-NOC",vars$source.description )  

unique(chem[pfchemexclude, "source.description"])
unique(vars[pfvars, "source.description"])

```

Textile dye
```{r}
#Note, only liquid laundry detergent was used in the Risk Evaluation
dychem=grepl("dye",chem$source.description ) 
dyvars=grepl("dye",vars$source.description ) 
chem[dychem,] #note, there is concentration data in sheds for dye
vars[dyvars,]

```

Spray Polyurethane foam
```{r}
sfchem=grepl("spray foam",chem$source.description ) 
sfvars=grepl("spray foam",vars$source.description ) 
chem[sfchem,] #note, there is concentration data in sheds for dye
vars[sfvars,]
```


Reduce vars file to only parameters necessary for model 
```{r}
params=c("mass", "use.freq", "use.prev", "f.drain")


```

Combine categories together. Note: the chem file is less useful here because 1,4D is noted here as resulting from byproducts. Then, add quantiles from the appropriate distribution 

```{r}
prodvarrow=data.frame(scvars, afvars, dsvars, dwvars, ldvars, pfvars, dyvars, sfvars)

#vars[prodvarrows[which(prodvarrows$scvars==TRUE), ],]
quants=c(0.01,seq(0.1,0.9,0.1),0.99)
vars$form=tolower(vars$form)
quantlist=list()
vars[vars$CV<1 & vars$form=="lognormal",] # it looks like none of the products that have CV < 1 are ones we need, but it bothers me to have NA's. 

#The problem with the solution here is that if a sd gets moved to 1, than under lognormal conditions, there is not variation since it goes to 0. 
for(i in 1:length(vars[,1])){
sub=vars[i,]
if(sub$form=="point" | sub$form=="uniform"){
  sub1=rep(sub$mean,length(quants))} else 
  {
   gm=sub$mean/sqrt(1+sub$CV^2)
   gsd=exp(sqrt(log(1+sub$CV^2)))
    sub1= qlnorm(quants, log(gm), log(gsd))}
  
quantlist=c(quantlist,list(sub1))} 
  
quantab=as.data.frame(do.call("rbind", quantlist))
names(quantab)=c(paste0("Q", c(0.01,seq(0.1,0.9,0.1),0.99)))

vars=data.frame(vars, quantab)


vars[vars$source.description=="dish soap",]


```



#Finally, compile only the variables we are interested in and write out all the excel files. 
```{r}
productlist=NULL
library(xlsx)
write.xlsx(vars,file="Consumer_Product_Vars_SHEDS_14D.xlsx", sheetName="Vars")
write.xlsx(prodvarrow, file="Consumer_Product_Vars_SHEDS_14D.xlsx", sheetName="Extracted", append=TRUE)
for(i in 1:length(prodvarrow[1,])){
sub=vars[which(prodvarrow[,i]==TRUE),]
sub=subset(sub, variable%in%params)
productlist=rbind(productlist, sub)}

#write.xlsx(productlist, file="Consumer_Product_Vars_SHEDS_14D.xlsx", sheetName = products[i], append=TRUE)}
write.xlsx(productlist, file="Con_Prod_Vars_SHEDS_14D_SingleSheet_DED120821.xlsx")

```

a log normal distribution with a mean of 5 and an sd of 1 means that if you take the mean of the log-transformed values of the distribution, it'll be 5. 

SHEDS takes the arithmetic mean and sd and log transforms them to sample from the log distribution. It makes sure that they aren't below 0 and 1 respectively, by transforming them to the geometric mean and sd.  

qlnorm(c(0.05,0.25,0.5,0.75,0.95), meanlog=log(1000), sdlog=log(100)) #Specifies the values of each quantile

plot(dlnorm(c(1:100), meanlog=log(10), sd=log(10) )) #@specifies the probability density at each value

plot(plnorm(c(1:100), meanlog=log(10), sd=1 )) #specifies the comulative probability density 

```

v$par1[f=="logn"] <- v$mean[f=="logn"] /(1+v$cv[f=="logn"]^2)^0.5
v$par2[f=="logn"] <- exp(log(1+v$cv[f=="logn"]^2)^0.5)




